taxadb: A High-Performance Local Taxonomic Database Interface

### 2019-04-01 paper/manuscript.Rmd
WIP

journal: "Methods in Ecology & Evolution"
date: "`r Sys.Date()`"
bibliography: refs.bib
output: rticles::elsevier_article
---

```{r message=FALSE, include = FALSE}
pkgconfig::set_config("dplyr::na_matches" = "never")
```

@@

Databases of taxonomic names such as the Integrated Taxonomic Information System [ITIS; @itis], the National Center for Biological Information's (NCBI) Taxonomy database, [@ncbi], the Catalogue of Life [COL; @col], and over one hundred other providers have sought to address these problems by providing expert-curated lists of accepted taxonomic names, synonyms, associated taxonomic rank, hierarchical classification, and scientific authority (e.g. author and date) establishing a scientific name.
The R language [@R] is widely used in ecology and evolution [@Lai2019] and the `taxize` package [@Chamberlain2012] has become a popular way R users to interact with naming providers and name resolution services.  `taxize` implements bindings to the web APIs (Application Programming Interface) hosted by many popular taxonomic name providers. Unfortunately, this means that functions in the `taxize` are impacted by several major drawbacks that are inherent in the implementation of these central API servers, such as:

@@

```{r}
library(taxadb)
```

@@

```{r}
trait_data <- read_tsv(system.file("extdata", "trait_data.tsv"))
status_data <- read_tsv(system.file("extdata", "status_data.tsv"))
```

```{r}
status_data
```

@@

by_name(top_5, "col") %>%
  select(scientificName, acceptedNameUsageID) %>% 
  distinct() %>% 
  mutate(acceptedNameUsage = get_names(acceptedNameUsageID, "col"))
```

```{r}
by_rank("Aves", rank="class", provider = "col")
```

@@

itis_accepted <- taxa_tbl("itis") %>%
  filter(taxonomicStatus == "accepted", taxonRank == "species") %>% 
  count() %>% pull(n)
```
`r itis_col` of the `r itis_accepted` of *accepted names* in ITIS (`r itis_col / itis_accepted * 100`%) are not recognized by the (2018) Catalogue of Life as either accepted names or known synonyms.

>>>>

journal: "Methods in Ecology & Evolution"
date: "`r Sys.Date()`"
bibliography: refs.bib
layout: 3p
output: rticles::elsevier_article
---

```{r message=FALSE, include = FALSE}

pkgconfig::set_config("dplyr::na_matches" = "never")

library(printr)
knit_print.data.frame <- function(x, ...) {
#  knitr::knit_print(DT::datatable(x), ...)
  knitr::knit_print(kableExtra::kable(x, "latex"), ...)
}

knitr::opts_chunk$set(cache=FALSE)
```

@@

Databases of taxonomic names such as the Integrated Taxonomic Information System [ITIS; @itis], the National Center for Biological Information's (NCBI) Taxonomy database, [@ncbi], the Catalogue of Life [COL; @col], and over one hundred other providers have sought to address these problems by providing expert-curated lists of accepted taxonomic names, synonyms, associated taxonomic rank, hierarchical classification, and scientific authority (e.g. author and date) establishing a scientific name.
The R language [@R] is widely used in ecology and evolution [@Lai2019] and the `taxize` package [@Chamberlain2013] has become a popular way R users to interact with naming providers and name resolution services.  `taxize` implements bindings to the web APIs (Application Programming Interface) hosted by many popular taxonomic name providers. Unfortunately, this means that functions in the `taxize` are impacted by several major drawbacks that are inherent in the implementation of these central API servers, such as:

@@

```{r message = FALSE, warning=FALSE}
library(tidyverse)
library(taxadb)
```

@@

```{r}
trait_data <- read_tsv(system.file("extdata", "trait_data.tsv", package="taxadb"))
status_data <- read_tsv(system.file("extdata", "status_data.tsv", package="taxadb"))
```

```{r }
status_data
```

@@

by_name(top_5, "col") %>%
  select(scientificName, acceptedNameUsageID) %>% 
  distinct() %>% 
  mutate(acceptedNameUsage = get_names(acceptedNameUsageID, "col")) %>% 
  head(5)
```

```{r}
by_rank("Aves", rank="class", provider = "col") %>% head(5)
```

@@

itis_accepted <- taxa_tbl("itis") %>%
  filter(taxonomicStatus == "accepted", taxonRank == "species") %>% 
  count() %>% pull(n)
```
$`r itis_col`$ of the $`r itis_accepted`$ of *accepted names* in ITIS ($`r itis_col / itis_accepted * 100`$%) are not recognized by the (2018) Catalogue of Life as either accepted names or known synonyms.
